### 2024-08-11 doc/Thesis/chapter/021-theory-calculations.tex
start doing todo.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Artificial Intelligence and Machine Learning}
%How can \gls{ai} and \gls{ml} be defined? 
\Gls{ai} is a trans-disciplinary field with roots in logic, statistics, cognitive psychology, decision theory, neuroscience, linguistics, cybernetics, and computer engineering\cite{howard2019artificial}.
The history of \gls{ai} goes back to the middle of the \nth{20} century. 
Researchers from the emerging field came together at a 1956 Dartmouth conference and the term "\gls{ai}" was coined\cite{McCarthy1955}. 
\Gls{ai}'s  history is beautifully depicted by McCorducks' 1982 book "Machines Who Think"\cite{McCorduck1982,Apter1982}, which focuses on the great minds behind the advances.
Pioneers like Alan Turing thought a lot about how to define, test and implement \gls{ai}\cite{howard2019artificial}. 
One example how to measure \gls{ai} is to let it play chess against a human\cite{Silver2017} 
(in 1997 a chess computer called Deep Blue won against the World Chess Champion Garry Kasparov for the first time\cite{Feng1999}).
Another test \textit{ingenioused} by Alan Turing is the imitation game\cite{turing1950imitation}, nowadays known as Turing test:
an interrogator communicates with two unknown entities \textbf{A} and \textbf{B} (a woman and a man) and must find out who is who. 
\textbf{A} will try to make to interrogator misjudge, whereas \textbf{B} is on the interrogators side.
The question is if \textbf{A} is replaced with a computer how the ratio of outcomes would deviate from the original ratio. 
%
At the moment it is hard to imagine a computer getting a higher ratio than a human
but when reading AI written articles\cite{gpt2020}, it's easy to see this test being passed in the near future; especially with ChatGPT's convincingly human-like answers\cite{dis2023chatgpt}.

But this does not mean that computers are more intelligent than humans or \linebreak[4] sentient\cite{searle1980,searle1999married} even though some claim\cite{tiku2022google} and it certainly does not mean that research is over. 
AI is still a young field, which is strongly growing and is gaining ubiquitous status. 
It is slowly creeping into every aspect of modern human life just like electricity around one hundred years ago. 
Realms in which AI is gaining traction are: 
%
playing board games (and beating humans)\cite{Silver2017,Feng1999,Campbell2002}, 
image recognition (very popular for medical diagnosis)\cite{Li2020,Deo2015,Topol2019,Fujiyoshi2019}, 
chemistry\cite{Westermayr2019,goh2017chemception,jha2018elemnet}, 
cyber security\cite{Sarker2021},
facial recognition (to prevent theft of lavatory paper at public toilet in China\cite{Andrews2017}),
financial sector (as robo-advisors)\cite{Littman2021},
natural language processing (NLP)\cite{Koroteev2021,Liu2021gpt,Parviainen2021} 
(which can also create code and pictures through scalable vector graphics (SVG))
and even creative tasks like 
creating non existing faces\cite{Mansourifar2020}, 
create graphic artwork (DALL-E 2)\cite{Marcus2022} or 
making video games\cite{Guzdial2016}.
%
It is nearly hard to find a field where \gls{ai} is not used in some way. 
This steady incorporation of \gls{ai} leads to the so called \gls{ai} effect\cite{McCorduck1982,ai100}: 
certain fields get incorporated into \gls{ai} research and practice,
such that after some time of general use it is no more considered \gls{ai} (e.g. spam filter or web searches).
Google CEO Sundar Pichai even goes as far and said: 
"AI is one of the most important things humanity is working on. It is more profound than [...] electricity or fire"\cite{Hassan2020}.

@@

\Gls{nn} (one of the most popular architectures for big data\cite{Chiroma2019}) are loosely modelled after the brain\cite{bishop1994neural}.
Artificial neurons (also called nodes), which are arranged in layers, 
are connected to each of the neurons of previous and next layers
and the weights (parameters of intensity), with which the data is routed from one neuron to another, 
are optimised during training. 
%
Convolutional-layer \gls{nn} excel in picture recognition\cite{Lecun1995conv} and are useful in quantum mechanics too\cite{westermayr2020combining}.
Other common methods include linear regression, kernel ridge regression and support vector regression.

@@

There are also lesser known \gls{ml} algorithms such as 
evolutionary algorithms (e.g. \gls{ga} and \gls{pso}).
Evolutionary algorithms take advantage of the ability to cope with local optima by evolving several candidate solutions simultaneously\cite{villanova2010function}.
%
%One advantage 
One particular feature of evolutionary algorithms is that they start with a small data set
and periodically request new data in order to solve the problem iteratively.
%\td{This is exactly what I needed because every experiment is very time intensive.}

@@

A \gls{pso} also uses a starting population of particles where each experiment (particle) 
is represented by its independent and dependent variables. 
It was originally inspired by the behavior of bird flocks and fish schools\cite{villanova2010function,Kennedy1995}.
%"PSO is a population based heuristic inspired by the flocking behavior of birds. 
%To simulate the behavior of a swarm, each bird (or particle) is allowed to fly towards the optimum solution."\cite{villanova2010function}
Each particle has an associated position and velocity. 
Every movement across the search space is, 
additionally to a stochastic term, 
%stochastic and 
influenced by its particle velocity and position as well as its and the swarm's best visited position.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{EMMA}
\subsection{Evolutionary Model-based Multiresponse Approach}
\Gls{emma} is an implementation of \gls{pso} in the \texttt{R} programming language. 
Each time step the dependent variables for all possible input variable combinations %$\mathbf{x} = \{x_1,\cdots ,x_N\}$ 
are predicted with the help of \gls{mars}. 
The \gls{mars} \gls{rf} is then used to chose the next position for each particle.
\Gls{mars} is a regression method introduced by Friedman's 1991 paper 
"Multivariate Adaptive Regression Splines"\cite{friedman1991multivariate}. 
The \texttt{CRAN} (Comprehensive R Archive Network) package which implements \gls{mars} is called \texttt{earth} due to \gls{mars} being trade marked\cite{mars}.
%%% MARS 
\gls{mars} is an extension of multivariate linear regression.
Friedman presented \gls{mars} as an alternative to piecewise polynomials (splines) and local averaging methods (e.g. kernel functions). %, whose number of parameter quickly \td{exceeds the practical} for moderate dimensional problems ($n>5$). 
Advantages of \Gls{mars} are that it uses less parameters and it produces continuous models with continuous derivatives.
%
%to higher dimensions (n > 2) is straightforward in principle but difficult in practice. 
%"These difficulties are related to the so-called “curse-of-dimensionality”, a phrase coined by Bellman (1961) to express the fact that exponentially increasing numbers of points are needed to densely populate Euclidean spaces of increasing dimension."\cite{friedman1988fitting}
%
Furthermore, \gls{mars} has the ability to fit (possibly complex) interactions and non-linearities without losing interpretability. 
%
%- "The key ingredient that advances this approach to general settings is the ability to fit (possibly complex) interactions among the variables through the product terms that are permitted to enter the approximation (9), if required by the fit."\cite{friedman1988fitting}
%% MARS MODEL SELECTION
%- "The approximation is developed in a forward/backwards stepwise recursive manner in analogy with the recursive partitioning approach. "\cite{friedman1988fitting}
%

@@

\iffalse
\subsection{Kernel Ridge Regression and Support Vector Machine}
\textbf{goal}
\begin{itemize}
    \item intro to stat and ml 
    \item say that although different will be abgehandelt in the same chapter 
    \item basic statistics: MAE vs MSE, R, std, F-test 
    \item lin reg 
    \item ridge regression 
    \item kernel ridge regression 
    \item support vector machine 
    \item anova (categorical), PCA, step wise regression
    \item all previous ananlyse data to predict, but don't tell me which data point is needed next 
    \item DOE (add data acquisition)
    \item genetic algorithm and pso (they combine data gen with regression/prediction and optimization, actually incorporate prediction into selection process but don't per se specify which pred and optimization)
    \item comment: always tradeoff between complexity and interpretability
    \item quesiton: which paragraph by GPT2?? 
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%% STATISTICS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Design of Experiments} %DOE
%\epigraph{"The real purpose of experiment design is to maximize the information content of the data within the limits imposed by the given constraints."}{Grahem C. Goodwin\cite{goodwin1977experiment}}
%"Our ability to design a good experiment should depend upon our prior knowledge regarding the nature of the data generating mechanism."\cite{goodwin1977experiment}
%
\begin{quote}
	{"The real purpose of experiment design is to maximize the information content of the data within the limits imposed by the given constraints."}
	- {Grahem C. Goodwin\cite{goodwin1977experiment}}
\end{quote}
%
In two cases a deliberate \gls{doe} is especially beneficial.
(1) If the query of a new data point is very expensive, it is favourable to actively chose the query (e.g. drilling for oil or quantum chemical calculations).

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Artificial Intelligence and Machine Learning}
%How can \gls{ai} and \gls{ml} be defined? 
\Gls{ai} is a trans-disciplinary field with roots in logic, statistics, cognitive psychology, decision theory, neuroscience, linguistics, cybernetics, and computer engineering\cite{howard2019artificial}.
The history of \gls{ai} goes back to the middle of the \nth{20} century. 
Researchers from the emerging field came together at a 1956 Dartmouth conference and the term "\gls{ai}" was coined\cite{McCarthy1955}. 
\Gls{ai}'s  history is beautifully depicted by McCorducks' 1982 book "Machines Who Think"\cite{McCorduck1982,Apter1982}, which focuses on the great minds behind the advances.
Pioneers like Alan Turing thought a lot about how to define, test and implement \gls{ai}\cite{howard2019artificial}. 
One example how to measure \gls{ai} is to let it play chess against a human\cite{Silver2017} 
(in 1997 a chess computer called Deep Blue won against the World Chess Champion Garry Kasparov for the first time\cite{Feng1999}).
Another test \textit{ingenioused} by Alan Turing is the imitation game\cite{turing1950imitation}, nowadays known as Turing test:
an interrogator communicates with two unknown entities \textbf{A} and \textbf{B} (a woman and a man) and must find out who is who. 
\textbf{A} will try to make to interrogator misjudge, whereas \textbf{B} is on the interrogators side.
The question is if \textbf{A} is replaced with a computer how the ratio of outcomes would deviate from the original ratio. 
%
At the moment it is hard to imagine a computer getting a higher ratio than a human
but when reading AI written articles\cite{gpt2020}, it is easy to see this test being passed in the near future; especially with ChatGPT's convincingly human-like answers\cite{dis2023chatgpt}.

But this does not mean that computers are more intelligent than humans or \linebreak[4] sentient\cite{searle1980,searle1999married} even though some claim\cite{tiku2022google} and it certainly does not mean that research is over. 
AI is still a young field, which is strongly growing and is gaining ubiquitous status. 
It is slowly creeping into every aspect of modern human life just like electricity around one hundred years ago. 
Realms in which AI is gaining traction are: 
%
playing board games (and beating humans)\cite{Silver2017,Feng1999,Campbell2002}, 
image recognition (very popular for medical diagnosis)\cite{Li2020,Deo2015,Topol2019,Fujiyoshi2019}, 
chemistry\cite{Westermayr2019,goh2017chemception,jha2018elemnet}, 
cyber security\cite{Sarker2021},
facial recognition (to prevent theft of lavatory paper at public toilet in China\cite{Andrews2017}),
financial sector (as robo-advisors)\cite{Littman2021},
natural language processing (NLP)\cite{Koroteev2021,Liu2021gpt,Parviainen2021} 
(which can also create code and pictures through scalable vector graphics (SVG))
and even creative tasks like 
creating non existing faces\cite{Mansourifar2020}, 
create graphic artwork (DALL-E 2)\cite{Marcus2022} or 
making video games\cite{Guzdial2016}.
%
It is nearly hard to find a field where \gls{ai} is not used in some way. 
This steady incorporation of \gls{ai} leads to the so called \gls{ai} effect\cite{McCorduck1982,ai100}: 
certain fields get incorporated into \gls{ai} research and practice,
such that after some time of general use it is no more considered \gls{ai} (e.g.\ spam filter or web searches).
Google CEO Sundar Pichai even goes as far and said: 
"AI is one of the most important things humanity is working on. It is more profound than [...] electricity or fire"\cite{Hassan2020}.

@@

\Gls{nn} (one of the most popular architectures for big data\cite{Chiroma2019}) are loosely modelled after the brain\cite{bishop1994neural}.
Artificial neurons (also called nodes), which are arranged in layers, 
are connected to each of the neurons of previous and next layers
and the weights (parameters of intensity), with which the data is routed from one neuron to another, 
are optimized during training. 
%
Convolutional-layer \gls{nn} excel in picture recognition\cite{Lecun1995conv} and are useful in quantum mechanics too\cite{westermayr2020combining}.
Other common methods include linear regression, kernel ridge regression and support vector regression.

@@

There are also lesser known \gls{ml} algorithms such as 
evolutionary algorithms (e.g.\ \gls{ga} and \gls{pso}).
Evolutionary algorithms take advantage of the ability to cope with local optima by evolving several candidate solutions simultaneously\cite{villanova2010function}.
%
%One advantage 
One particular feature of evolutionary algorithms is that they start with a small data set
and periodically request new data in order to solve the problem iteratively.
%\td{This is exactly what I needed because every experiment is very time intensive.}

@@

A \gls{pso} also uses a starting population of particles where each experiment (particle) 
is represented by its independent and dependent variables. 
It was originally inspired by the behavior of bird flocks and fish schools\cite{villanova2010function,Kennedy1995}.
%"PSO is a population based heuristic inspired by the flocking behavior of birds. 
%To simulate the behavior of a swarm, each bird (or particle) is allowed to fly towards the optimum solution."\cite{villanova2010function}
Each particle has an associated position and speed. 
Every movement across the search space is, 
additionally to a stochastic term, 
%stochastic and 
influenced by its particle speed and position as well as its and the swarm's best visited position.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{EMMA}
\subsection{Evolutionary Model-based Multiresponse Approach}
\Gls{emma} is an implementation of \gls{pso} in the \texttt{R} programming language. 
Each time step the dependent variables for all possible input variable combinations %$\mathbf{x} = \{x_1,\cdots ,x_N\}$ 
are predicted with the help of \gls{mars}. 
The \gls{mars} \gls{rf} is then used to chose the next position for each particle.
\Gls{mars} is a regression method introduced by Friedman's 1991 paper 
"Multivariate Adaptive Regression Splines"\cite{friedman1991multivariate}. 
The \texttt{CRAN} (Comprehensive R Archive Network) package which implements \gls{mars} is called \texttt{earth} due to \gls{mars} being trade marked\cite{mars}.
%%% MARS 
\gls{mars} is an extension of multivariate linear regression.
Friedman presented \gls{mars} as an alternative to piecewise polynomials (splines) and local averaging methods (e.g.\ kernel functions). %, whose number of parameter quickly \td{exceeds the practical} for moderate dimensional problems ($n>5$). 
Advantages of \Gls{mars} are that it uses less parameters and it produces continuous models with continuous derivatives.
%
%to higher dimensions (n > 2) is straightforward in principle but difficult in practice. 
%"These difficulties are related to the so-called “curse-of-dimensionality”, a phrase coined by Bellman (1961) to express the fact that exponentially increasing numbers of points are needed to densely populate Euclidean spaces of increasing dimension."\cite{friedman1988fitting}
%
Furthermore, \gls{mars} has the ability to fit (possibly complex) interactions and non-linearities without losing interpretability. 
%
%- "The key ingredient that advances this approach to general settings is the ability to fit (possibly complex) interactions among the variables through the product terms that are permitted to enter the approximation (9), if required by the fit."\cite{friedman1988fitting}
%% MARS MODEL SELECTION
%- "The approximation is developed in a forward/backwards stepwise recursive manner in analogy with the recursive partitioning approach. "\cite{friedman1988fitting}
%

@@

\iffalse
\subsection{Kernel Ridge Regression and Support Vector Machine}
\textbf{goal}
\begin{itemize}
    \item intro to stat and ml 
    \item say that although different will be abgehandelt in the same chapter 
    \item basic statistics: MAE vs MSE, R, std, F-test 
    \item lin reg 
    \item ridge regression 
    \item kernel ridge regression 
    \item support vector machine 
    \item anova (categorical), PCA, step wise regression
    \item all previous ananlyse data to predict, but dont tell me which data point is needed next 
    \item DOE (add data acquisition)
    \item genetic algorithm and pso (they combine data gen with regression/prediction and optimization, actually incorporate prediction into selection process but dont per se specify which pred and optimization)
    \item comment: always tradeoff between complexity and interpretability
    \item quesiton: which paragraph by GPT2?? 
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%% STATISTICS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Design of Experiments} %DOE
%\epigraph{"The real purpose of experiment design is to maximize the information content of the data within the limits imposed by the given constraints."}{Grahem C. Goodwin\cite{goodwin1977experiment}}
%"Our ability to design a good experiment should depend upon our prior knowledge regarding the nature of the data generating mechanism."\cite{goodwin1977experiment}
%
\begin{quote}
	{"The real purpose of experiment design is to maximize the information content of the data within the limits imposed by the given constraints."}
	- {Grahem C. Goodwin\cite{goodwin1977experiment}}
\end{quote}
%
In two cases a deliberate \gls{doe} is especially beneficial.
(1) If the query of a new data point is very expensive, it is favourable to actively chose the query (e.g.\ drilling for oil or quantum chemical calculations).

### 2024-08-11 doc/Thesis/chapter/022-alt-theory-calculations.tex
start doing todo.md

\subsection{Introduction to AI, ML and statistics}
\textbf{What is AI?} It is the interdisciplinary field of making a artificial inteligent agent. 
What is intelligence though?
Intelligence is not easily defined. Some definitions I agree with is that intelligence 
includes learning from experience, finding solutions to new problems. 
Creativity is also substencial for intelligence. 
It is a goal in the field AI to immitate (and surpass) human intelligence. 
The turing tests aims to test a computers intelligience by testing if it can trick a person into believing that the computer is person while a person tries to make the person believe that they are in fact a person. 
Approaches range from knowledge based systems over hard coded intstructions to machine learning. 
%
\textbf{Machine Learning} is the art of telling a machine or computer how to gain knowlegde from data (aka experience). 
There are different problems and solutions. 
The problems include image recognition, voice recognition, natural language processing, data prediction, classification, recommendation systems and so on. 
%
In order to solve these problems machine learning resorts to statistical methods such as
linear regression, ridge regression. 
Further machine learning methods not regularly seen in statistics are the 
kernel trick, support vector machines, genetic algorithms and particle swarm algorithms. 
The last two are special in the sense that they don't provide a certain regression method,
but rather a frame work to choose the next sample which should be tested in a optimization. 
\td{mention classification vs regression and supervized vs supervized}
%
\textbf{Statistics} is a subfield of mathematics and emerged from the aim to summerzie large data sets 
and extract information. In contrast to descriptive statistic, there is stochastic. 
Stochastic statistic aims on predicting data points with probabilities. 
%
\textbf{Linear regression} is one of the simplest methods to predict data. 
It besticht by its computational simplicity and easy interpretation. 
%Let's take this equation
\begin{equation}
\mathbf{y} = \mathbf{X} \mathbf{k} +c 
\end{equation}
Where $\mathbf{y}$ (size $\mathbb{R}^n$, where $n$ is the number of data points) is a vector of dependent variables and $\mathbf{X}$ (size $\mathbb{R}^{n\times d}$, where $d$ is the the number of independent variables) is a matrix of independent variables. 
The parameters $\mathbf{k}$ (size $\mathbb{R}^d$) and $c$ (size $\mathbb{R}$) are chosen by minimizing an objective function (loss function in \gls{ml} chargon).
A typical function to minimize are the L1 and L2 error, i.e. \gls{mae} and \gls{mse}, respectively.
\begin{align}
L_1&= \sum |y_i - \hat y_i| = \sum \sum |y_i - x_{i,j}\cdot k_i+c| \\
L_2&= \sum(y_i - \hat y_i)^2 = \sum (y_i - x_{i,j}\cdot k_i+c)^2
\end{align}
%

@@

\begin{itemize}
    \item \sout{intro to stat and ml}
    \item \sout{say that although different will be abgehandelt in the same chapter}
    \item \sout{basic statistics: MAE vs MSE, R, std, F-test} 
    \item \sout{lin reg}
    \item \sout{ridge regression}
    \item \sout{kernel ridge regression}
    \item \sout{support vector machine}
    \item anova (categorical), PCA, step wise regression
    \item all previous ananlyse data to predict, but don't tell me which data point is needed next 
    \item DOE (add data acquisition)
    \item genetic algorithm and pso (they combine data gen with regression/prediction and optimization, actually incorporate prediction into selection process but don't per se specify which pred and optimization)
    \item comment: always tradeoff between complexity and interpretability
    \item quesiton: which paragraph by GPT2?? 
\end{itemize}

>>>>

\subsection{Introduction to AI, ML and statistics}
\textbf{What is AI?} It is the interdisciplinary field of making a artificial inteligent agent. 
What is intelligence though?
Intelligence is not easily defined. Some definitions I agree with is that intelligence 
includes learning from experience, finding solutions to new problems. 
Creativity is also substencial for intelligence. 
It is a goal in the field AI to immitate (and surpass) human intelligence. 
The turing tests aims to test a computers intelligience by testing if it can trick a person into believing that the computer is person while a person tries to make the person believe that they are in fact a person. 
Approaches range from knowledge based systems over hard coded intstructions to machine learning. 
%
\textbf{Machine Learning} is the art of telling a machine or computer how to gain knowlegde from data (aka experience). 
There are different problems and solutions. 
The problems include image recognition, voice recognition, natural language processing, data prediction, classification, recommendation systems and so on. 
%
In order to solve these problems machine learning resorts to statistical methods such as
linear regression, ridge regression. 
Further machine learning methods not regularly seen in statistics are the 
kernel trick, support vector machines, genetic algorithms and particle swarm algorithms. 
The last two are special in the sense that they dont provide a certain regression method,
but rather a frame work to choose the next sample which should be tested in a optimization. 
\td{mention classification vs regression and supervized vs supervized}
%
\textbf{Statistics} is a subfield of mathematics and emerged from the aim to summerzie large data sets 
and extract information. In contrast to descriptive statistic, there is stochastic. 
Stochastic statistic aims on predicting data points with probabilities. 
%
\textbf{Linear regression} is one of the simplest methods to predict data. 
It besticht by its computational simplicity and easy interpretation. 
%Let's take this equation
\begin{equation}
\mathbf{y} = \mathbf{X} \mathbf{k} +c 
\end{equation}
Where $\mathbf{y}$ (size $\mathbb{R}^n$, where $n$ is the number of data points) is a vector of dependent variables and $\mathbf{X}$ (size $\mathbb{R}^{n\times d}$, where $d$ is the the number of independent variables) is a matrix of independent variables. 
The parameters $\mathbf{k}$ (size $\mathbb{R}^d$) and $c$ (size $\mathbb{R}$) are chosen by minimizing an objective function (loss function in \gls{ml} chargon).
A typical function to minimize are the L1 and L2 error, i.e.\ \gls{mae} and \gls{mse}, respectively.
\begin{align}
L_1&= \sum |y_i - \hat y_i| = \sum \sum |y_i - x_{i,j}\cdot k_i+c| \\
L_2&= \sum(y_i - \hat y_i)^2 = \sum (y_i - x_{i,j}\cdot k_i+c)^2
\end{align}
%

@@

\begin{itemize}
    \item \sout{intro to stat and ml}
    \item \sout{say that although different will be abgehandelt in the same chapter}
    \item \sout{basic statistics: MAE vs MSE, R, std, F-test} 
    \item \sout{lin reg}
    \item \sout{ridge regression}
    \item \sout{kernel ridge regression}
    \item \sout{support vector machine}
    \item anova (categorical), PCA, step wise regression
    \item all previous ananlyse data to predict, but dont tell me which data point is needed next 
    \item DOE (add data acquisition)
    \item genetic algorithm and pso (they combine data gen with regression/prediction and optimization, actually incorporate prediction into selection process but dont per se specify which pred and optimization)
    \item comment: always tradeoff between complexity and interpretability
    \item quesiton: which paragraph by GPT2?? 
\end{itemize}
